### 2001-05-28 technical/printing/printing.tex
Initial import: Representations for printing
svn/trunk@47

%
% Natural Language Toolkit for Python
% Representations for Printing
% Edward Loper
% 
% Created [05/27/01 11:44 PM]
% $Id$
%

\documentclass[11pt]{article}
\usepackage{fullpage}

\begin{document}
\title{Representations for Printing\\
\Large Natural Language Processing Toolkit: Technical Report}
\author{Edward Loper}
\maketitle

%============= INTRODUCTION =============%
\section{Introduction}

\end{document}