### 2001-12-07 technical/printing/printing.tex
- Revised design goals - Added section discussing str vs repr, etc. - Added a list of guidelines (incomplete)
svn/trunk@403

% Natural Language Toolkit Technical Report:
% Representations for Printing
%
% Copyright (C) 2001 University of Pennsylvania
% Author: Edward Loper <edloper@gradient.cis.upenn.edu>
% URL: <http://nltk.sf.net>
% For license information, see LICENSE.TXT
%
% $Id$

\documentclass[11pt]{article}
\usepackage{fullpage}

\begin{document}
\title{Representations for Printing\\
\Large Natural Language Processing Toolkit: Technical Report}
\author{Edward Loper}
\maketitle

% One thing that guides peoples intuitions about what an object is is
% how it looks when it prints. If two things look the same when you
% print them, then intuitively they should be the same. I think this may
% have been a source of confusion in the first problem set, where Tokens
% look just like strings when you print them, so users will be tempted
% to do things like use == to test whether a TokenType is equal to a
% string.

%============= INTRODUCTION =============%
\section{Introduction}

  The printed representations of an object guides users' intuitions
  about the content, function, and purpose of the object.  Often, a
  user's first encounter with an object is with its printed
  representation.  Users rely on this representation to let them know
  what type of object it is, and what its important features are.  The
  printed representations of each type of object defined by the NLP
  toolkit must be carefully considered.

%============= GOALS =============%
\section{Goals}

  The following goals guide the design of printed representations.

  \subsection{Primary Goals}

  \begin{description}

    \item[Type Indication] The representation should give some direct
    or indirect indication of its type.  In particular, it should be
    easy to tell the type of an object from its printed
    representation.

    \item[Intuitiveness] Representations should rely on the
    conventional meanings of formats and symbols (e.g., the use of
    ``@'' to denote ``at'').  This will make it easier for users to
    understand what the representations stand for.

    \item[Standard Representations] When possible, established
    standard representations (such as ``dog/NP'' for tagging) should
    be used.

    \item[Consistancy] The printed representation of an object should
    be consistant between different environments.  Inconsistant
    representations can produce confusion about object values and
    object types.

    \item[Composition] The representation of an object should indicate
    what data fields it abstractly contains.  In particular, for each
    data field abstractly contained by an object, its representation
    should include some indication that the field is contained within
    it.

  \end{description}

  \subsection{Secondary Goals}

  \begin{description}

    \item[Value Indication] If two objects have the same value, they
    should have the same printed representations; if two objects have
    different values, then they should have different printed
    representations.

    \item[Compactness] Printed representations should be simple and
    compact.  Simpler representations are easier to read and easier to
    understand, especially when the object is included in the
    representation of a larger data structure, such as a list of
    objects.

  \end{description}

%============= PYTHON =============%
\section{Python Constructs}

  Python classes may define two different string representations: one
  is returned by \texttt{str}, and the other by \texttt{repr}.  The
  differences between these two represenetations are:

  \begin{itemize}

    \item The ``Official'' Definitions
    (\texttt{http://www.python.org/doc/current/ref/customization.html})
    \begin{itemize}
      \item 
        \texttt{repr}: If at all possible, this should look like a
        valid Python expression that could be used to recreate an
        object with the same value (given an appropriate
        environment). If this is not possible, a string of the form
        "$<$...some useful description...$>$" should be returned.
        This is typically used for debugging, so it is important that
        the representation is information-rich and unambiguous.

      \item
        \texttt{str}: This differs from \_\_repr\_\_() in that it does not
        have to be a valid Python expression: a more convenient or
        concise representation may be used instead. 
    \end{itemize}

    \item Usage Contexts
    \begin{itemize}

      \item \texttt{repr}
      \begin{itemize}
        \item The result of \texttt{repr()}
        \item The result of backquotes (e.g., \texttt{`x`})
        \item The result of the string formatting operator
            ``\texttt{\%r}''
        \item Used when an object representation is ``contained'' in
            another object representation (e.g., to display an element
            of a list or a key of a dictionary)
      \end{itemize}

      \item \texttt{str}
      \begin{itemize}
        \item The result of \texttt{str()}
        \item The result of the string formatting operator 
            ``\texttt{\%s}''
        \item The output of print
      \end{itemize}

    \end{itemize}

    \item Builtin Type Usage Conventions
    \begin{itemize}
      \item \texttt{repr(x)} contains no newlines.
      \item \texttt{repr(x)} contains only simple ASCII characters.
    \end{itemize}

    \item Default Values
    \begin{itemize}
      \item The default value for \texttt{repr} is a string of the form:\\
        \texttt{$<$\_\_\textit{module}\_\_.\textit{class} instance at \textit{address}$>$}
      \item The default value for \texttt{str} is the value returned
        by \texttt{repr}
    \end{itemize}

  \end{itemize}

  \subsection{The Official Definition}

    After consideration of the contexts in which \texttt{repr} and
    \texttt{str} are used, we have decided \emph{not} to make use of
    the ``official'' definitions.  This decision is primarily
    motivated by the fact that \texttt{repr} is always used when an
    object representation is ``contained'' in another object
    representation.  We believe that this would lead to overly verbose
    and difficult to read representations of complex objects.  For
    example, if we adopted the ``official'' definition of
    \texttt{repr}, then the string representation (using \texttt{repr}
    \emph{or} \texttt{str}) of a list of tokens would look something
    like:

\begin{verbatim}
"[Token('The', Location(0, 1, unit='w', source='foo.txt')), Token('dog', Lo
cation(1, 2, unit='w', source='foo.txt')), Token('saw', Location(2, 3, unit
='w', source='foo.txt')), Token('the', Location(3, 4, unit='w', source='foo
.txt')), Token('man', Location(4, 5, unit='w', source='foo.txt'))]"
\end{verbatim}

    We believe that this is signifigantly less readable than the
    result of using a \texttt{repr} which does not evaluate to return
    its original value:

\begin{verbatim}
"['The'@[0w]@'foo.txt', 'dog'@[1w]@'foo.txt', 'saw'@[2w]@'foo.txt', 'the'@[
3w]@'foo.txt', 'man'@[4w]@'foo.txt']"
\end{verbatim}

  \subsection{str vs repr}

    Instead, we will make the following distinction between the
    \texttt{str} and \texttt{repr} string representations:

    \begin{itemize}

      \item \texttt{repr} gives a single-line representation, which
          would be suitable for inclusion in the representation of
          another object (i.e., it is delimited by quotes, brackets,
          braces, etc., when appropriate).  This representation should
          contain only standard ASCII characters.  \texttt{repr}
          should be defined for all classes.

      \item \texttt{str} can optionally give a more human-readable
          representation of an object.  This representation may span
          multiple lines.
    \end{itemize}

\section{Guidelines}

  The following guidelines are intended to help design string
  representations for new nltk objects.  In these guidelines, the
  string representation of an object is refered to as \texttt{s}.

  \subsection{General Guidelines}
    \begin{itemize}
      \item All classes should implement \texttt{repr}.  When the
        output of \texttt{repr} may be difficult to read, \texttt{str}
        should be implemented.
      \item If a field has no value (or the default value), it may be
        ommited from the printed representation.  (E.g.,
        \texttt{Location}s with no source don't display the source.
      \item When \texttt{s} needs to contain the representation of an
        embedded object, and the object's type is variable,
        \texttt{repr} should always be used (not \texttt{str}).  No
        special type-checking should be applied (e.g., to treat
        strings specially).
      \item It should always be possible to tell the type of an object
          from its printed representation.
    \end{itemize}

  \subsection{Repr}
    \begin{itemize}
      \item \texttt{s} may not span multiple lines (i.e., it may not
        contain newlines)
    \end{itemize}

  \subsection{Str}
    \begin{itemize}
      \item \texttt{s} may span multiple lines.
      \item Multiline \texttt{str} representations shoud not contain any
        trailing newlines.
      \item \texttt{s} may use indentation to show internal structure.
      \item \texttt{s} may be a simple table.
      \item If \texttt{s} is a multiline representation, it should try 
        very hard not to be more than 75 characters wide.
      \item It should always be possible to distinguish the
          \texttt{s} from the \texttt{str} for any built-in object
          (except strings, which can have arbitrary representations);
          or for any other class.
      \item For multiline \texttt{str} representations, \texttt{s}
        should begin with an unindented line, whose first word is the
        name of the class.  All subsequent lines should be indented.
    \end{itemize}

  \subsection{Data-Oriented Classes}
    \begin{itemize}
      \item When practical, \texttt{s} should specify the complete
          state of the object.  In particular, if it is possible to
          give the complete state of a typical instance in less than
          75 characters, then \texttt{s} should give the complete
          state.  
      \item When it is not practical to show the complete state
          of a data-oriented instance, this should be indicated with
          elipsis dots (``...'').
    \end{itemize}

  \subsection{Task-Oriented Classes}
    \begin{itemize}
      \item For repr, \texttt{s} should have the form:\\
        \texttt{$<$\textit{class} \textit{info}$>$}\\
        where \texttt{\textit{info}} is a description of the
        parameters and the state of the class.
    \end{itemize}

%============= PP =============%
\section{Pretty-Printing Methods}

  Classes may provide pretty-printing methods.  These methods should
  have the following signature:

  \texttt{pp(self, left, right)}

  Where \texttt{left} and \texttt{right} are left and right margins.
  \texttt{pp} should assume that it is already at position
  \texttt{left}.  E.g., a pp function to draw a vertical line, when
  given \texttt{left}=5, would return:

  \texttt{"|" + ("~~~~~|"*lineheight)}

%============= EXAMPLES =============%
\section{Examples}

\subsection{Locations}

  \texttt{str} and \texttt{repr} representations are identical.  Unit
  \& Loc are not displayed if they are not specified.  End index is not
  displayed if it has the default value (=start+1).

{\tt
\begin{tabbing}
\emph{$>>>$ Location(1, 3)}\\
@[1:3]\\
\emph{$>>>$ Location(1, 2)}\\
@[1]\\
\emph{$>>>$ Location(1, 3, source='foo.txt')}\\
@[1:3]@'foo.txt'\\
\emph{$>>>$ Location(1, 3, unit='w')}\\
@[1w:3w]\\
\emph{$>>>$ Location(1, 2, unit='w', source='foo.txt')}\\
@[1w]@'foo.txt'\\
\emph{$>>>$ Location(1, 3, unit='w', source='foo.txt')}\\
@[1w:3w]@'foo.txt'
\end{tabbing}
}

\subsection{Tokens}

  \texttt{str} and \texttt{repr} representations are identical.
  Location is included using \texttt{repr}.  If location is
  unspecified, it is displayed as \texttt{@[?]}.  (Note that this
  would never be a valid location by itself.)

{\tt
\begin{tabbing}
\emph{$>>>$ Token('x', Location(1, 3, unit='w', source='foo.txt')}\\
'x'@[1w:3w]@'foo.txt'\\
\emph{$>>>$ Token('x')}\\
'x'@[?]
\end{tabbing}
}

\subsection{Trees}

  Lack of commas, and presence of ':', distinguises trees from tuples.

{\tt
\begin{tabbing}
\emph{$>>>$ Tree('np', 'the', 'boy')}\\
('np': 'the' 'boy') \\
\emph{$>>>$ Tree('vp', 'likes', Tree('np', 'the', 'boy'))}\\
('vp': 'likes' ('np': 'the' 'boy'))\\
\emph{$>>>$ TreeToken('vp', Token('likes',1), TreeToken('np', Token('John',2)))}\\
('vp': 'likes' ('np': 'John))@[1:3]\\
\\
\textrm{Trees have one-line repr representations, \& multiline str representations:}\\
\emph{$>>>$ Tree('x', 'a'*40, Tree('b', 'c'*40, Tree('d', 'e'*40)))}\\
('x': 'aaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaa' ('b': 'cccccccccccccccccccccccc\\
cccccccccccccccc' ('d': 'eeeeeeeeeeeeeeeeeeeeeeeeeeeeeeeeeeeeeeee')))\\
\emph{$>>>$ print Tree('x', 'a'*40, Tree('b', 'c'*40, Tree('d', 'e'*40)))}\\
('x':\\
\qquad  'aaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaa'\\
\qquad  ('b':\\
\qquad\qquad    'cccccccccccccccccccccccccccccccccccccccc'\\
\qquad\qquad    ('d': 'eeeeeeeeeeeeeeeeeeeeeeeeeeeeeeeeeeeeeeee')))
\end{tabbing}
}

\subsection{SimpleFreqDists}

{\tt
\begin{tabbing}
\emph{$>>>$ s = SimpleFreqDist(\ldots)}\\
\{SimpleFreqDist $<$CFSample 1, '7'$>$:1, $<$CFSample 5, '2'$>$:1, $<$CFSample 3, '2'$>$\\
:1, $<$CFSample 2, '7'$>$:1, $<$CFSample 7, '2'$>$:2, $<$CFSample 9, '7'$>$:2, $<$CFSampl\\
e 6, '7'$>$:3, $<$CFSample 5, '7'$>$:1, $<$CFSample 3, '7'$>$:1, $<$CFSample 9, '9'$>$:2,\\
 $<$CFSample 7, '9'$>$:3, $<$CFSample 8, '5'$>$:5\}\\
\emph{$>>>$ print s}\\
\{SimpleFreqDist\\
\qquad $<$CFSample 1, '7'$>$: 1, $<$CFSample 5, '2'$>$: 1, $<$CFSample 3, '2'$>$: 1, \\
\qquad $<$CFSample 2, '7'$>$: 1, $<$CFSample 7, '2'$>$: 2, $<$CFSample 9, '7'$>$: 2, \\
\qquad $<$CFSample 6, '7'$>$: 3, $<$CFSample 5, '7'$>$: 1, $<$CFSample 3, '7'$>$: 1, \\
\qquad $<$CFSample 9, '9'$>$: 2, $<$CFSample 7, '9'$>$: 3, $<$CFSample 8, '5'$>$: 5\}
\end{tabbing}
}

%============= ISSUES =============%
\section{Issues}

The representations are still under development.  This section
describes some of the issues involved in designing representations.  I
may not use the solutions discussed here.

\subsection{Context-Dependant Representations}

  Representations that show the complete state of an object are useful
  for certain tasks, but they can be overly verbose for complex
  objects, or for lists of objects, making it hard to understand the
  structures shown by the printed representations.  In particular,
  complete representations of complex objects are not \emph{compact}.

  The principle of \emph{composition} requires that the representation
  display each field of an object; but it may be useful to display a
  ``summary'' of some of the fields, instead of a complete
  representation.  For example, a \texttt{Location} whose unit is
  \texttt{"token"} and whose source is \texttt{"foo.txt"} might
  normally be represented as:

  \begin{quote}
    \texttt{@[token 1:token 3]@foo.txt}
  \end{quote}

  But if a large list of tokens is displayed, each with a location,
  then this representation may be overly verbose.  A ``summary'' of
  this location might be:

  \begin{quote}
    \texttt{@[1:3]}
  \end{quote}

  Note that this solution is somewhat contrary to \emph{consistancy}:
  we are using different representations for the same object in
  different environments.  However, if summary representations are
  chosen carefully enough, I don't believe that this will be a
  problem.

\subsubsection{\texttt{str} and \texttt{repr}}

  The \texttt{str} and \texttt{repr} functions could be used to
  implement the difference between full representations and summary
  representations.  Officially, \texttt{str} gives the ``informal
  representation'' of an object and \texttt{repr} gives its ``formal
  representation.''  However, one of the main differences between them
  in practice is that \texttt{str} is used when a single object is
  printed, but \texttt{repr} is used when an object is printed as part
  of a list, tuple, or dictionary.

  We could extend this distinction to the toolkit's objects, using
  \texttt{str} to encode an object's top-level representation, and
  \texttt{repr} to encode its summary.  The advantage of this approach
  is that it guarantees that each object has exactly one summary
  representation and one full representation.  For some objects, these
  may be the same representation.

\subsection{Refactoring Information}

  Unfortunately, summary representations don't always seem like the
  right tool to make representations more compact.  In a number of
  cases, the representation of a complex object may contain a large
  amount of redundant information.  It may make sense to display this
  information exactly once.  e.g., syntax trees: how to attatch
  locations?

%============= RANDOM NOTES =============%
\section{Random Notes}

\subsection{Issues}

There are a number of considerations when deciding how to represent
objects:

\begin{itemize}

  \item Users should be able to tell what type of object something is
    by printing it.  In particular:
    \begin{itemize}
      \item It should be possible to distinguish an object from any
        built-in type (string, list, etc)
      \item It should be possible to distinguish each of the different
        classes defined by the toolkit from each other.
    \end{itemize}

    Of course, this is within reason -- a string can look like any
    object, for example.

  \item How much to display?  Representations that show the complete
  state of an object can be useful, but they can be overly verbose for
  large objects, or for lists of objects, making it hard to understand
  the structures shown by the printed representations.

  \item One rep per object, or not?  In particular, should there be a
  different embedded representation?  Difference between str and repr
  as embedding.

  \item Representations should be intuitive

  \item Representations should be simple

  \item Should representations be readable?

  \item Should standard reps be used (eg, treebank)

  \item Should some classes be treated specially (eg, quote strings)

\end{itemize}

\end{document}