### 2003-08-02 technical/conventions/conventions.tex
- Expanded on various sections.
svn/trunk@1032

% Natural Language Toolkit Technical Report:
% Conventions
%
% Copyright (C) 2001 University of Pennsylvania
% Author: Edward Loper <edloper@gradient.cis.upenn.edu>
% URL: <http://nltk.sf.net>
% For license information, see LICENSE.TXT
%
% $Id$

\documentclass{article}
\usepackage{fullpage}
\usepackage{parskip}
\setlength{\parindent}{0ex}
\setlength{\parskip}{1ex}

\begin{document}
\title{Conventions}
\author{Edward Loper}
\maketitle

%################################################################
%#  Overview
%################################################################
\section{Overview}

This document lists design and coding conventions that are used
throughout NLTK.  Whenever possible, these conventions should be
followed.

You should also look at the source code of existing NLTK modules, to
get a feel for conventions that aren't explicitly documented here.

%################################################################
%#  Module Requirements
%################################################################
\section{Module Requirements}

The following checklist lists requirements for a module to be added to
NLTK.  (Some of these requirements may be fulfilled after the module
is added to NLTK, but they should be fulfilled eventually.)

\begin{itemize}
\item Reference documentation
  \begin{itemize}
  \item The module is documented by a docstring, describing the
    purpose of the module, and describing how the objects defined by
    the class relate to that purpose.
  \item Each interface is documented by a docstring, describing the
    requirements imposed by the interface.
  \item Each class is documented by a docstring, describing the class.
  \item Each function or method is documented by a docstring,
    including documentation for parameters, parameter types, return
    value, and return value type.
  \end{itemize}
\item Tutorial documentation
  \begin{itemize}
  \item Provide motivation for the module's task.
  \item Explain how to use the module to perform that task.
  \item Explain how to create and manipulate any data structures
    defined by the module.
  \item Explain any algorithms used.
  \end{itemize}

\item Unit testing code
  \begin{itemize}
  \item Each function or method is tested
  \end{itemize}
\end{itemize}

%################################################################
%#  Mutability
%################################################################
\section{Immutable vs Mutable Variables}

When possible, basic data types (such as \texttt{Token},
\texttt{Tree}, and \texttt{CFGProduction}) should be immutable.  An
\textit{immutable} object is one whose value cannot be changed, once
the object is created.  In Python, integers and strings are immutable;
but lists and dictionaries are mutable.

Immutable objects have several useful properties, but for our
purposes, the most important are:

\begin{itemize}
\item They can be used as types for tokens.
\item They can be used as keys in dictionaries.
\item They can be stored in Sets (\texttt{nltk.set.Set}).
\end{itemize}

The following guidelines should help you decide whether an object
should be mutable or immutable:

\begin{itemize}
\item Any class that represents a text type \textit{must} be
  immutable.
\item Classes that represent fixed values, such as CFG productions or
  rules for chunking, should be immutable.
\item Container classes (such as parsing charts) are usually mutable.
\item If there's not a good reason to make a class mutable, then keep
  it immutable.
\end{itemize}

If you're unsure about whether a class should be mutable or immutable,
contact \texttt{edloper@gradient.cis.upenn.edu}.

\subsection{Making a Class Immutable}

To make a class immutable:
\begin{itemize}
\item Do not define any methods that modify its contents.
\item Do not define any methods that return a reference to a mutable
  instance variable.  (E.g., don't return a list that's used
  internally; otherwise, the caller might modify it).  If you need to
  return a mutable value, then copy it.  But you should also consider
  using immutable structures (e.g. tuples) for your instance
  variables.
\item Define the special method \texttt{\_\_cmp\_\_}, and optionally
  define the two special methods \texttt{\_\_eq\_\_} and \texttt{\_\_ne\_\_}.
\item Define the special method \texttt{\_\_hash\_\_}.
\end{itemize}

%################################################################
%#  Naming Objects
%################################################################
\section{Naming Objects}

Object names should be descriptive, but not overly long.

\subsection{Variable Names}

Name variables with lower case names.  Use underscores to separate
words in the variable name.  Some conventional names for common types
are:

\begin{tabular}{l|l}
Type & Conventional Names \\
\hline
\texttt{Token} & \texttt{token}, \texttt{tok}\\
\texttt{Type} & \texttt{type}, \texttt{typ}, \texttt{word}\\
\texttt{FreqDist} & \texttt{freqdist}, \texttt{fdist}\\
\texttt{ProbDistI} & \texttt{probdist}, \texttt{pdist}\\
\texttt{ConditionalFreqDist} & \texttt{cfreqdist}, \texttt{cfdist}\\
\texttt{ConditionalProbDistI} & \texttt{cprobdist}, \texttt{cpdist}\\
\texttt{TreeToken} & \texttt{treetok}\\
\texttt{Tree} & \texttt{tree}\\
\texttt{CFGProduction} & \texttt{prod}\\
\texttt{aggedType} & \texttt{ttype}\\
\end{tabular}

These names are often prefixed by a descriptor, such as
``\texttt{context\_token}''.

\subsection{Class Names}

Name classes with WordInitialCaps.  Name interfaces with a trailing
``I'' (e.g. \texttt{ProbDistI} or \texttt{ParserI}).

\subsection{Module Names}

Name modules with short lower-case names.  Modules that define data
types are generally named after the primary data types they define
(e.g., \texttt{nltk.cfg} and \texttt{nltk.token}).  Modules that
define NLP tasks are generally named after the task that they define
(e.g., \texttt{nltk.tagger} and  \texttt{nltk.parser}).

%################################################################
%#  Interfaces
%################################################################
\section{Interfaces}

An \textit{interface} is a special type of base class that specifies a
set of methods that must be supported.  Each subclass of an interface
provides an implementation for the methods defined by the interface.

Interfaces are implemented as classes, all of whose methods raise
exceptions.  Required methods should raise \texttt{AssertionError}s;
and optional methods should raise \texttt{NotImplementedError}s.  Each
interface method defines a docstring that describes the behavior of
that method.  See existing interfaces for examples.

%################################################################
%#  String Representations
%################################################################
\section{String Representations}

\begin{itemize}
\item \texttt{repr} should return a single-line description of an
  object, suitable for embedding in other descriptions.

\item \texttt{str} should produce a description that is suitable for
  the output of \texttt{print}.  The \texttt{str} description may span
  multiple lines, and does not need to have clear delimiters.
\end{itemize}

%################################################################
%#  Task Interfaces
%################################################################
\section{Task Interfaces \& Classes}

\subsection{Multiple Return Values}

\subsection{Trace Output}

trace output is controlled by a single ``trace level'' instance
variable.

trace values: 0: no output; 1: less than a page for a typical run; 2
or higher: more verbose.

setting trace output:
      - by default, trace=0
      - constructor takes a trace keyword argument
      - trace(n) method sets trace level; use default value for n.

\end{document}