### 2003-10-14 technical/interfaces/interfaces.tex
initial import
svn/trunk@1348

% Natural Language Toolkit Technical Report:
% Interfaces
%
% Copyright (C) 2003 University of Pennsylvania
% Author: Edward Loper <edloper@gradient.cis.upenn.edu>
% URL: <http://nltk.sf.net>
% For license information, see LICENSE.TXT
%
% $Id$

\documentclass[11pt]{article}
\usepackage{fullpage, alltt}

\begin{document}
\title{Interfaces\\
\Large Natural Language Processing Toolkit: Technical Report}
\author{Edward Loper}
\maketitle

%============= INTRODUCTION =============%
\section{Introduction}

An \textit{interface} is a specification of the behavior for a set of
related methods.  NLTK defines standard interfaces to ensure that
different components will work together; and to allow different
implementations of the same component to be easily substituted for
each other.  NLTK defines two types of interface:

\begin{itemize}
\item \textit{Processing interfaces} are used to define natural
  language processing tasks (such as parsing or classifying texts).
\item \textit{Data type interfaces} are used to define standard
  methods for accessing data types that have multiple implementations
  (such as probability distributions).
\end{itemize}

%============= BACKGROUND =============%
\section{Background}
\subsection{Protocols}

Python does not provide any language-level support for defining or
enforcing interfaces.  Instead, Python tends to rely on
\textit{protocols}, which are informal interfaces that are specified
by external documentation.  Protocols are typically defined relative
to existing data types or classes.  For example, the ``sequence''
protocol requires that an object implement the methods that tuples and
lists implement.\footnote{Or some subset of these methods; the exact
  requirements are not usually made explicit.}

This informal interface system allows Python functions to be used
polymorphically: you can call a function with a object type that it
doesn't expect, as long as you guarantee that the object behaves
correctly.  It also allows individual functions to only require
certain aspects of an protocol.  For example, a function that reads from
a file-like object might just require that its input implement the
\texttt{read} and \texttt{readline} methods.

\newpage\noindent
However, Python's informal protocol system has several disadvantages:
\begin{itemize}
\item In order to avoid duplicating documentation, interfaces must be
  documented externally (not in docstrings).  However, there is no
  standard place to doccument interfaces; and since the interfaces are
  separated from the code, it is easy for the interfaces to get
  out-of-sync with the code.
\item Since protocols are usually documented informally, different
  programmers may make different assumptions about what is required or
  guaranteed.  These differences in assumption can lead to subtle
  bugs, which can only be detected at run-time.
\item It is not possible to automatically check that classes correctly
  implement interfaces.
\item If a function requires only certain aspects of an protocol, then
  it can be more difficult to change that function's implementation.
  In the example above, it would be impossible to change the function
  that required a file-like object to an implementation that uses
  \texttt{seek} (without checking all calls to that function).
\end{itemize}

% Add other related work (e.g., zope ExtensionClasses?)

%============= INTERFACE CLASSES =============%
\section{Interface Classes}

For NLTK, we wanted to provide explicit formal definitions for all
interfaces.  We decided to encode these formal definitions using
special superclasses which we call ``interface classes.''  An
\textit{interface class} is a class that defines simple stubs for each
method that the interface specifies.  These stubs use docstrings to
describe the method's required behavior; and raise exceptions if
called.  This system has three primary advantages over external
documentation:

\begin{itemize}
\item For classes that implement an interface, the API documentation
  for methods can be inherited from the interface.\footnote{This
    assumes the use of an API documentation tool that supports
    documentation inheritance, such as Epydoc.}  This helps avoid
  duplicating documentation.
\item It is possible to automatically verify that classes define all
  of the required methods, using the required signatures.
\item It is easier to maintain consistancy between the interfaces'
  definitions and implementations in different versions of the
  toolkit.
\end{itemize}

\subsection{Naming Interface Classes}

Interface classes are always named with a trailing capital ``I,'' such
as \texttt{TokenizerI} or \texttt{EventI}.

\subsection{Interface Class Docstrings}

An interface class's docstring should describe the purpose of the
interface, and provide any necessary background information.  It
should also specify which methods (if any) are optional.

\subsection{Interface Class Constructors}

% Will this cause problems with the new super construct?
Interface classes should not define constructors.

\subsection{Interface Methods}

Every interface class defines a set of \textit{interface methods},
each of which specifies the behavior for a single method of the same
name.  An interface method consists of a declaration, followed by a
body containing a docstring and a \texttt{raise AssertionError(...)}
command:

\begin{alltt}
    def \textit{name}(\textit{parameters...}):
        """
        \textit{docstring}
        """
        raise AssertionError("\textit{name} not overridden")
\end{alltt}

\subsubsection{Optional Methods}

Interfaces may define ``optional methods,'' which should be
implemented when possible, but are not required.  The optionality of a
method is specified in two ways:

\begin{itemize}
\item Optional methods should raise a \texttt{NotImplementedError}
  exception (instead of an \texttt{AssertionError} exception).
\item The interface class docstring should provide a list of all
  optional methods.
\end{itemize}

\noindent
The following template should be used to write optional methods:

\begin{alltt}
    def \textit{name}(\textit{parameters...}):
        """
        \textit{docstring}
        """
        raise NotImplementedError("\textit{name} is not implemented")
\end{alltt}

\subsection{Implementing an Interface}

If a class implements an interface, then it must list that interface
as one of its base classes; and it must override each of the
interface's (required) methods.  Each overriding method must have a
compatible signature with the interface method\footnote{An overriding
  method is compatible with a base method if it will accept any set of
  parameters that the base method would accept.  For example, an
  overriding method may add a new optional argument; but it may not
  remove or rename an argument.}; and must have behavior that conforms
to the behavior described by the interface method's docstring.

Note that a single class can implement multiple interfaces, by
declaring them in its list of base classes, and overriding all of
their methods.

\subsection{Interface Inheritance Hierarchies}

Specialized interfaces can be derived from more general interfaces
using inheritance.  In particular, you can create a specialized
interface class by declaring it as a subclass of a more general
interface.  The documentation for any methods defined by the more
general interface will be inherited by the specialized interface.

For example, the \texttt{ParserI} interface class defines a general
interface for parsing; and its subclass \texttt{ProbabilisticParserI}
defines a more specialized interface for probabilistic parsing.

%============= TYPES OF INTERFACE =============%
\section{Types of Interface}

NLTK defines a large set of interfaces.  To help organize these
interfaces, we have categorized them according to a number of
attributes.  The most basic division is between \textit{processing
  interfaces}, which are used to define natural language processing
tasks (such as parsing or classifying texts); and \textit{data type
  interfaces}, which are used to define standard interfaces to data
types that have multiple implementations.

%============= PROCESSING INTERFACES =============%
\section{Processing Interfaces}

The primary use of interfaces in NLTK is to define standard
specifications for natural language processing tasks.  Each processing
interface is named after a single processing task, such as
\texttt{TokenizerI} (for tokenizing) and \texttt{ParserI} (for
parsing).\footnote{A processing interface's name should be based the
  deverbal noun formed from the task's verb.}

Each processing interface defines at least one method, which performs
the language processing task and returns the result.  If no result is
found, then this method should return \texttt{None}.  The name for
this method should be based on imperitive form of the interface name,
such as \texttt{tokenize} (for tokenizing) and \texttt{parse} (for
parsing).

When appropriate, each processing interface should also define a
method that returns a list of the possible results for the given task,
sorted in descending order of quality.\footnote{For some
  implementation-specific definition of quality; note that it is
  perfectly acceptable to define all possible results to have the same
  quality, in which case results will be returned in arbitrary order.}
This method is named by appending ``\texttt{\_n}'' to the imperitive
form of the interface name, such as \texttt{tokenize\_n} for
tokenizing; and \texttt{parse\_n} for parsing.  The signature for this
methods should be formed by adding an optional parameter \texttt{n} to
the signature of the basic language processing task method, which
controls the maximum number of results that should be returned.  If
\texttt{n} is not specified, then all available results should be
returned.

\subsection{Input and Output}

Almost all processing interfaces define tasks that take tokens or
collections of tokens as input; and return tokens or collections of
tokens as outputs.  Some advantages of working with tokens rather than
types include:

\begin{itemize}
\item It makes it easy to keep track of what information was derived
  from what source; and to connect together pieces of information that
  have been generated by different tasks.
\item It makes it easy to differentiate two occurances of the same
  type.
\item It often aids in the evaluation of system performance, by
  allowing direct token-by-token comparison of two algorithms' output
  without the need for any alignment algorithms.
\end{itemize}

A notable exception to the rule that most interfaces take tokens as
inputs is the \texttt{tokenizer} method, which is responsible for
creating the original list of tokens from a string.

\newpage
Processing interfaces can be divided into classes, based on their
input and output.  Currently, all NLTK processing classes fit into the
following categories:

\begin{itemize}
\item \textit{Token list generator}: creates a list of tokens.
  \texttt{TokenizerI} is an example of a token list generator.
\item \textit{Token map}: takes a single token as its input, and
  returns a token with the same location and a new type.
  \texttt{ClassifierI} is an example of a token map.
\item \textit{Token list map}: takes a list of tokens as its input,
  and returns a list of tokens with the same locations, and new types.
  \texttt{TaggerI} is an example of a token list map.
\item \textit{Token list collector}: takes a list of tokens as its
  input, and returns a single new token composed from the list.
  \texttt{ParserI} is an example of a token list collector.
\item \textit{Processing class factory}: creates new instances of
  processing classes.  C{ClassifierTrainerI} is an example of a
  processing class factory.
\end{itemize}

Additional classes are likely to be added in the future.  Some likely
candidates include the following classes:

\begin{itemize}
\item \textit{Token list filter}: takes a list of tokens as its
  input, and returns a list containing a subset of the tokens from the
  input list (preserving order).  A stop-word list might be defined
  with this class of processing interface.
\item \textit{Token expander}: takes a single token as its input, and
  returns a list of tokens derived from the single token.  A
  morphological analyzer might be defined with this class of
  processing interface.
\end{itemize}

\subsection{Function}

Processing interfaces can be also be divided into classes based on
their basic function.  This provides a partially orthoganal
classification to the the input/output categories.  Currently, all
NLTK processing classes fit into the following categories:

\begin{itemize}
\item \textit{Extender}: Returns data containing a superset of the
  information in the input.  An example of a token extender is
  \texttt{TaggerI}, which adds tags to each tokens' types; or
  \texttt{ParserI}, which adds constituancy information.
\item \textit{Selector}: Returns data containing a subset of the
  information in the input.  An example of a selector is a stop-word
  list, which filters out tokens with a given set of types.
\item \textit{Transformer}: Returns data that contains neither a
  superset nor a subset of the information in the input.  An example
  of a transformer is an automatic speech recognition processing
  interface.
\end{itemize}

Note that the distinction between these classes is often \textit{not}
determined by the processing task itself.  For example, we could
define a tagging interface to return either a list of tagged words (an
extender) or a list of tags (a transformer).

%============= DATA TYPE INTERFACES =============%
\section{Data type Interfaces}

Data type interfaces are used to provide specifications for data types
that are too general to be captured by a single class.  For example,
the \texttt{ProbDistI} interface class defines an interface for
probability distributions, which can be implemented using a wide
variety of techniques.

In NLTK, data type interfaces are not nearly as common as processing
interfaces.  Because they introduce extra complexity, they should only
be used to define a data type when it is not practical to use a single
class.  Currently, NLTK defines data type interfaces for two domains:
probability distributions (\texttt{ProbDistI} and
\texttt{ConditionalProbDistI}); and classification features
(\texttt{FeatureDetectorI}, \texttt{FeatureDetectorListI}, and
\texttt{FeatureValueListI}).  The classification feature data type
interfaces may be removed or revised when the classification system is
redesigned.

%============= DEFINING NEW INTERFACES =============%
\section{Defining New Interfaces}

When defining a new interface for NLTK, be sure to keep the following
questions and guidelines in mind:

\begin{itemize}
\item Is it a data interface or a processing interface?  Note that in
  some cases, this will be a design decision (e.g., a ``rule'' class
  could be reformulated as a ``decider'' class).
\item For processing interfaces, which input/output class does the
  interface belong to?
\item For processing interfaces, which function class does the
  interface belong to?
  \begin{itemize}
  \item If you're trying to decide whether a processing interface
    should be an extender or a transformer, think about typical uses
    of the interface's output.  Is the new information usually used in
    conjunction with the original information, or in place of it?
  \end{itemize}
\item What auxilliary data types are needed to define the interface?
\item Interfaces should be as general as is practical, and should not
  introduce any unnecessary restrictions on tasks.  For example,
  interfaces should try not to constrain what objects can be used for
  token types.
\item Interfaces should be named using standard terms, where possible.
  Interface names (including interface method names) should not
  conflict with other names that are already used by the toolkit.
  Neutral names are preferred over domain-specific names.
\item Think about what tasks are likely to be used to generate input
  for the interface, or to process its output; and make sure that the
  interface can be easily combined with the interfaces for those other
  tasks.
\item Is the new interface just a special case of an existing
  interface?  For example, there is probably no need for a ``word
  sense disambiguation'' interface, since word sense disambiguation is
  just a specific domain for classification.
\item Is the nomenclature used to define the interface internally
  consistant?  Is it consistant with the literature?\footnote{An
    example of inconsistant nomenclature that is currently in NLTK is
    the fact that classifiers assign ``labels'' instead of
    ``classifications.''}

\end{itemize}

%============= CURRENT INTERFACES =============%
\section{Current Interfaces}
\subsection{Processing Interfaces}
The following table lists the processing interfaces that are currently
defined by NLTK, and lists the I/O and functional classifications for
each interface.

\vspace{3mm}\noindent
\begin{tabular}{|p{.5\textwidth}|l|l|}
\hline
  \textbf{Interface} & \textbf{I/O Classification} &
  \textbf{Function} \\
\hline
\texttt{TokenizerI}

\textit{\qquad Divides a text into a list of tokens}
& token list generator & transformer \\
\hline
\texttt{TaggerI}

\textit{\qquad Adds a tag to each token in a list}
& token list map & extender \\
\hline
\texttt{StemmerI}

\textit{\qquad Strips morphological affixes}
& token list map & selector \\
\hline
\texttt{ParserI}

\textit{\qquad Identifies token constituants}
& token list collector & extender \\
\hline
\texttt{ChunkParserI}

\textit{\qquad Identifies non-overlapping groups of tokens}
& token list collector & extender \\
\hline
\texttt{ProbabilisticParserI}

\textit{\qquad Identifies token constituants}
& token list collector & extender \\
\hline
\texttt{ClassifierI}

\textit{\qquad Adds a label to a text}
& token list map & extender \\
\hline
\texttt{ClassifierTrainerI}

\textit{\qquad Creates a new classifier}
& processing class factory & transformer \\
% Leave this out for now: \texttt{FeatureSelectorI} & 
\hline
\texttt{CorpusReaderI}

\textit{\qquad Parses and reads standard corpora}
& generator & (special)\\
\hline
\end{tabular}

\subsection{Data Interfaces}

The following list desribes the data interfaces that are currently
defined by NLTK.

\begin{enumerate}
\item \textbf{Probability Distributions}
  \begin{itemize}
  \item \textbf{\texttt{ProbDistI}} is an interface for probability
    distributions.  Probability distributions are typically derived
    from frequency distributions (e.g., \texttt{MLEProbDist}), but can
    also be created analytically (e.g., \texttt{UniformProbDist}).
  \item \textbf{\texttt{ConditionalProbDistI}} is an interface for
    conditional probability distributions.  Probability distributions
    that are derived from frequency distributions are typically
    created using the \texttt{ConditionalProbDist} implementation; but
    implementations for analytic distributions (eg gaussians) may also
    be added.
  \end{itemize}
\newpage
\item \textbf{Chart Parsers}
  \begin{itemize}
  \item \textbf{\texttt{ChartRuleI}} is an interface for rules that
    controls what edges should be added to a chart in a
    \texttt{ChartParser}.
  \item \textbf{\texttt{IncrementalChartRuleI}} is an interface for
    rules that controls what edges should be added to a chart in an
    \texttt{IncrementalChartParser}.
  \item \textbf{\texttt{EdgeI}} is an interface for edges in a chart.
    This class is used as a superclass for \texttt{ProductionEdge} and
    \texttt{TokenEdge}.\footnote{Since it is very unlikely to ever get
      other implementations, it might make more sense to change this
      class from an interface to an abstract superclass.}
  \end{itemize}
\item \textbf{Classificaiton Features}\footnote{The classification
    system is being redesigned, and these interfaces are likely to be
    removed.}
  \begin{itemize}
  \item \textbf{\texttt{FeatureDetectorI}} is an interface for
    ``feature detectors,'' which map labeled texts to values.
  \item \textbf{\texttt{FeatureValueListI}} is an interface for sparse
    encodings of lists of feature values.
  \item \textbf{\texttt{FeatureDetectorListI}} is an interface for
    sparse encodings of lists of feature detectors.
  \end{itemize}
\end{enumerate}

\section{Open Questions}

\subsection{Extenders vs Transformers}

For a given task, how should we decide whether its interface should be
an extender or a transformer?  Some examples of the alternatives for
current interfaces include:

\vspace{3mm}
\begin{tabular}{|l|ll|}
  \hline
  Task & Extender Interface Output& Transformer Interface Output\\
  \hline
  Tagging &
    list of Token($\langle$word, tag$\rangle$, loc) &
    list of Token(tag, loc) \\
  Classifying & 
    Token($\langle$text, label$\rangle$, loc) &
    Token(label, loc) \\
  Stemming & 
    Token($\langle$word, stem$\rangle$, loc) &
    Token(stem, loc) \\
  \hline
\end{tabular}
\vspace{3mm}

Note that the output would always be a token; at issue is what
information would be included in the token's type.  Note also that
an extender can be converted to a transformer by discarding the
input word/text; and that a transformer can be converted to an
extender by aligning its output with its input (using locations),
and ``gluing'' the input word/text to the output.

Currently, we're using a mix of extenders and transformers, and
attempting to choose the class that will most likely be useful to
users.  However, the decision between extenders and transformers can
be fairly arbitrary in some cases, and might cause confusion.

\newpage
One alternative would be to eliminate extenders, and make everything
a transformer.  The user would ``glue'' information from different
stages together, where necessary.  E.g., a user who wanted to tag
some text and send it through a lexicalized parser could run:

\begin{alltt}
  >>> toks = my\_tokenizer.tokenize(text)
  >>> tags = my\_tagger.tag(toks)
  >>> tagged\_toks = [Token(TaggedType(tok.type(), tag.type()), tok.loc())
  ...                for (tok, tag) in zip(toks, tags)]
  >>> tree = my\_parser.parse(tagged\_toks)
\end{alltt}

Another option would be to eliminate transformers, and make
everything an extender.  The user would discard unnecessary
information, if they don't want it.  For example, a user who wanted
to tag and parse text with an unlexicalized parser could run:

\begin{alltt}
  >>> toks = my\_tokenizer.tokenize(text)
  >>> tagged\_toks = my\_tagger.tag(toks)
  >>> tags = [Token(tok.type().tag(), tok.loc()) for tok in tagged\_toks]
  >>> tree = my\_parser.parse(tags)
  >>> lexicalized\_tree = tree.lexicalize(toks)\footnote{This function\
is hypothetical.}
\end{alltt}

So, should we continue to use a mix of transformers and extenders?  Or
should we use one or the other exclusively?

\subsection{Accessing Context}

One issue with the token-based interfaces is that it can be difficult
to access a token's context.  For example, to disambiguate the sense
of a word, we might want to know what words are near it; but if a
processing class is only given the token, then it can't access this
information.

There are several ways we might address this issue, but each has its
own problems.

\begin{itemize}
\item Add pointers from tokens to the texts that they came from.
  Disadvantages: maintaining back pointers can be dangerous and error
  prone.  For example, if the user modifies the text, then the back
  pointers may become invalid.  This could potentially be solved by
  making all texts immutable (e.g., use tuples instead of lists for
  tokenized text); but that seems somewhat draconian.

\item Add an explicit feature-collection step, which acts on complete
  texts, and maps each token to a set of features that describes the
  token (e.g., words in its context).  Interfaces that need access to
  context would then act on these feature tokens, rather than on the
  original tokens.  Disadvantages: added complexity can be complex.
  Sometimes ``context'' is not best thought of as a set of features.

\item Add an extra ``text'' parameter to most interfaces, which can be
  used to look up the context of a token.  Disadvantages: finding the
  context for a token is likely to require linear search, and so would
  be very slow; interfaces become less simple and clean.
\end{itemize}

\end{document}