### 2007-07-12 slides/baypiggies.tex
incomplete slides on RTE
svn/trunk@4871

\documentclass{beamer}             % for slides
% \documentclass[handout]{beamer}    % for handout
\input{beamer}

\title{Introduction to Natural Language Processing}

\author{Steven Bird \and Ewan Klein \and Edward Loper}
\institute{
  University of Melbourne, AUSTRALIA
  \and
  University of Edinburgh, UK
  \and
   University of Pennsylvania, USA
}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{What is NLTK?}

\begin{frame}
  \titlepage
\end{frame}

\subsection{Introduction}

\begin{frame}
  \frametitle{Introduction}
  \begin{itemize}
    \item What is NLP?
    \item What is NLTK?
    \item ...
    \item ...
  \end{itemize}
\end{frame}

\subsection{Data Formats}

\begin{frame}
  \frametitle{Data Formats}
  \begin{itemize}
    \item ...
    \item ...
    \item ...
    \item ...
    \item ...
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Overview of Talk}
  \begin{enumerate}
    \item ...
    \item ...
    \item ...
    \item ...
    \item ...
  \end{enumerate}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{What is NLP?}

\begin{frame}
  \frametitle{NLP}

\begin{block}{Major sub-domains of NLP}
  \begin{itemize}
    \item Understanding
    \item Translation
    \item Generation
  \end{itemize}
\end{block}

\begin{block}{Modalities}
  \begin{itemize}
    \item Text
    \item Speech
    \item Images
  \end{itemize}
\end{block}

\end{frame}

\begin{frame}
  \frametitle{Textual Inference}

\begin{exampleblock}{Example 1}
  \begin{description}
    \item [Text] Never before had ski racing, a sport dominated by
      monosyllabic mountain men, seen the likes of Alberto Tomba, the
      flamboyant Bolognese flatlander who at 21 captured two gold
      medals at the Calgary Olympics.
    \item [Hypothesis] Alberto Tomba won a race.
  \end{description}
\end{exampleblock}

\end{frame}

\begin{frame}
  \frametitle{Textual Inference}

\begin{exampleblock}{Example 2}
  \begin{description}
    \item [Text] Researchers at the Harvard School of Public Health
      say that people who drink coffee may be doing a lot more than
      keeping themselves awake---this kind of consumption apparently
      also can help reduce the risk of diseases.
    \item [Hypothesis] Coffee drinking has health benefits.
  \end{description}
\end{exampleblock}

\end{frame}

\begin{frame}
  \frametitle{Why is it useful?}

  \begin{description}
    \item[Information Extraction] 
    \item[Spoken Dialogue] 
    \item[Question Answering] Who won races at the Calgary Olympics?
    \item[Text Summarization] 
  \end{description}

\end{frame}

\subsection{Notes}

\begin{frame}
  \frametitle{Notes}
  \begin{itemize}
    \item QA, Summarization, Dialog (NL understanding; holy grail of NLP)
    \item RTE
    \item Low-level components
    \item ambiguity, why NLP is hard, PP attachment, combinatorial explosion
    \item limitations of NLP, avoiding AI-completeness
    \item statistical vs rule-based
    \item limitations of just throwing more data at a problem
  \end{itemize}
\end{frame}

\pgfdeclareimage[width=8cm]{tag-context}{../images/tag-context}
\begin{frame}
  \frametitle{Example of an Image}
  \pgfuseimage{tag-context}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Case Studies}

\subsection{??}

\begin{frame}
  \frametitle{Notes}
  \begin{itemize}
    \item Text classification
    \item Distributional similarity
    \item Parsing
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Getting Involved}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}