### 2007-07-12 slides/baypiggies.tex
added intro slides
svn/trunk@4873

\documentclass{beamer}             % for slides
% \documentclass[handout]{beamer}    % for handout
\input{beamer}

\title{Introduction to Natural Language Processing}

\author{Steven Bird \and Ewan Klein \and Edward Loper}
\institute{
  University of Melbourne, AUSTRALIA
  \and
  University of Edinburgh, UK
  \and
   University of Pennsylvania, USA
}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{What is NLTK?}

\begin{frame}
  \titlepage
\end{frame}

\subsection{Introduction}

\begin{frame}
  \frametitle{Natural Language Processing}
  \begin{itemize}
    \item text and information
      \note[item]{world is awash with text; information locked up in text}
      \note[item]{sources: news, blogs, transcribed speech, other languages}
    \item Natural Language Processing (NLP)
      \note[item]{automatically process text, extracting structure and content, to give people access to the information}
    \item Engineering goals
      \note[build a system to provide a service (translation, summarization)]
    \item Scientific goals
      \note[analyze the structure of language, computational linguistics]
    \item Data
  \end{itemize}
\end{frame}

\subsection{Data Formats}

\begin{frame}[fragile]
  \frametitle{Chunked Tagged WSJ Text}
{\scriptsize\begin{verbatim}
[ Pierre/NNP Vinken/NNP ]
,/,
[ 61/CD years/NNS ]
old/JJ ,/, will/MD join/VB
[ the/DT board/NN ]
as/IN
[ a/DT nonexecutive/JJ director/NN Nov./NNP 29/CD ]
./.

[ Mr./NNP Vinken/NNP ]
is/VBZ
[ chairman/NN ]
of/IN
[ Elsevier/NNP N.V./NNP ]
,/,
[ the/DT Dutch/NNP publishing/VBG group/NN ]
./.
\end{verbatim}}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Parsed WSJ Text}
{\scriptsize\begin{verbatim}
( (S
    (NP-SBJ
      (NP (NNP Pierre) (NNP Vinken) )
      (, ,)
      (ADJP
        (NP (CD 61) (NNS years) )
        (JJ old) )
      (, ,) )
    (VP (MD will)
      (VP (VB join)
        (NP (DT the) (NN board) )
        (PP-CLR (IN as)
          (NP (DT a) (JJ nonexecutive) (NN director) ))
        (NP-TMP (NNP Nov.) (CD 29) )))
    (. .) ))
\end{verbatim}}
\end{frame}

\begin{frame}[fragile]
  \frametitle{IOB-Tagged Text (I=inside, O=outside, B=begin)}
{\scriptsize\begin{verbatim}
Characteristically RB B-ADVP
, , O
he PRP B-NP
did VBD B-VP
n't RB I-VP
waste VB I-VP
much JJ B-NP
time NN I-NP
getting VBG B-VP
started VBN I-VP
. . O
\end{verbatim}}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Named-Entity Classification}
{\tiny\begin{verbatim}
<b_enamex type="LOCATION">NAIROBI<e_enamex>, <b_enamex type="LOCATION">Kenya<e_enamex>
(<b_enamex type="ORGANIZATION">AP<e_enamex>) _ <b_numex type="CARDINAL">Thousands<e_numex>
of laborers, students and opposition politicians on <b_timex type="DATE">Saturday<e_timex>
protested tax hikes imposed by their cash-strapped government, which they accused of
failing to provide basic services.  Beneath a scorching sun, they sang anti-government
songs and chanted "<b_enamex type="PERSON">Moi<e_enamex> must go," showing their derision
for President <b_enamex type="PERSON">Daniel arap Moi<e_enamex>,
<b_enamex type="LOCATION">Kenya<e_enamex>'s ruler for <b_timex type="DURATION">20 years<e_timex>.
By voice vote, the <b_numex type="CARDINAL">5,000<e_numex> protesters approved a resolution
calling for the government to scrap new taxes, convene a convention to write a new
Constitution, stop harassing students and street vendors, and halt ethnic violence.
\end{verbatim}}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Tagged Biomedical Text in XML}
{\tiny\begin{verbatim}
<sentence id="@@0001"> <interactor pos="NN">Coexpression</interactor>
<token pos="IN">of</token> <token pos="NN">mouse</token> <gene>PKR ( 1-515 ) WT</gene>
<token pos="IN">as</token> <token pos="DT">a</token> <gene>Gal4</gene>
<token pos="VBG">DNA-binding</token> <token pos="NN">domain</token>
<token pos="NN">fusion</token> <token pos="IN">with</token>
<token pos="RB">either</token> <token pos="DT">the</token>
<token pos="JJ">catalytic-deficient</token> <token pos="JJ">human</token>
<gene>PKR ( 1-551 )</gene> <token pos="IN">K296R</token>
<token pos="JJ">mutant</token> <token pos=",">,</token> <token pos="DT">the</token>
<token pos="NNP">RNA-binding-deficient</token> <token pos="JJ">human</token>
<gene>PKR ( 1-551 ) K64E / K296R</gene> <token pos="RB">double</token>
<token pos="JJ">mutant</token> <token pos=",">,</token> <token pos="CC">or</token>
<token pos="JJ">wild-type</token> <token pos="NN">mouse</token>
<gene>PKR ( 1-515 ) WT</gene> <token pos="IN">as</token>
<token pos="JJ">full-length</token> <gene>PKR-Gal4</gene>
<interactor pos="NN">activation</interactor> <token pos="NN">domain</token>
<token pos="NNS">fusions</token> <token pos="VBD">resulted</token>
<token pos="IN">in</token> <interactor pos="NN">activation</interactor>
<token pos="IN">of</token> <token pos="DT">the</token> <gene>HIS3</gene>
<token pos="CC">and</token> <gene>lacZ</gene> <token pos="NNS">reporters</token>
<token pos=".">.</token></sentence>
\end{verbatim}}
\end{frame}

\begin{frame}[fragile]
  \frametitle{CMU Pronunciation Lexicon}
{\scriptsize\begin{verbatim}
APPROVE 1 AH0 P R UW1 V
APPROVED 1 AH0 P R UW1 V D
APPROVES 1 AH0 P R UW1 V Z
APPROVING 1 AH0 P R UW1 V IH0 NG
APPROVINGLY 1 AH0 P R UW1 V IH0 NG L IY0
APPROXIMATE 1 AH0 P R AA1 K S AH0 M AH0 T
APPROXIMATE 2 AH0 P R AA1 K S AH0 M EY2 T
APPROXIMATED 1 AH0 P R AA1 K S AH0 M EY2 T AH0 D
APPROXIMATELY 1 AH0 P R AA1 K S AH0 M AH0 T L IY0
APPROXIMATES 1 AH0 P R AA1 K S AH0 M AH0 T S
APPROXIMATES 2 AH0 P R AA1 K S AH0 M EY2 T S
APPROXIMATING 1 AH0 P R AA1 K S AH0 M EY2 T IH0 NG
APPROXIMATION 1 AH0 P R AA2 K S AH0 M EY1 SH AH0 N
APPROXIMATIONS 1 AH0 P R AA2 K S AH0 M EY1 SH AH0 N Z
\end{verbatim}}
\end{frame}

\subsection{NLTK}

\begin{frame}
  \frametitle{Natural Language Toolkit}
  \begin{enumerate}
    \item facilitate access to data
      \note[item]{only write a format reader once, then share}
      \note[item]{reading a format correctly is harder than you think, cf csv}
    \item standard algorithms
      \note[item]{getting into NLP requires knowledge of linguistics, CS, AI which is hard to get}
      \note[item]{clarity of Python: people can read the reference implementation, try it, reimplement it...}
    \item tutorials
      \note[item]{major textbooks are formidable, encyclopedic}
      \note[item]{we've selected useful core, and provided practical learning materials (book, demos, source)}
      \note[item]{practical: learning-by-doing; theoretically sound: stepping stone to research literature}
    \item contents
      \note[item]{50k lines of code, 300Mb datasets, 380pp book}
    \item evolution: beginnings, redesign
      \note[item]{teaching NLP, got too heavy using nested dictionaries, people ended up "programming in NLTK", rather than programming in Python using NLTK where needed}
      \note[item]{NLTK-Lite: completely reimplemented, simpler structure, uses new language features like set, generator expressions, defaultdict, elementtree}
      \note[item]{Design philosophy: interface definitions, lightweight reference implementations, utilities for common tasks}
  \end{enumerate}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Find 20 Most Frequent Words in a Text}

{\scriptsize\begin{verbatim}
>>> counts = defaultdict(int)
>>> for word in text.split():
...     counts[word] += 1
>>> ranked_words = sorted(counts.items, key=itemgetter(1))
>>> print [word for (word,count) in reversed(ranked_words)][:20]
\end{verbatim}}

\note{defaultdict; sorting tuples, list comprehensions}

{\scriptsize\begin{verbatim}
>>> print FreqDist(tokenize.wordpunct(text)).sorted()[:20]
\end{verbatim}}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{What is NLP?}

\begin{frame}
  \frametitle{NLP}

\begin{block}{Major sub-domains of NLP}
  \begin{itemize}
    \item Understanding
    \item Translation
    \item Generation
  \end{itemize}
\end{block}

\begin{block}{Modalities}
  \begin{itemize}
    \item Text
    \item Speech
    \item Images
  \end{itemize}
\end{block}

\end{frame}

\begin{frame}
  \frametitle{Textual Inference}

\begin{exampleblock}{Example 1}
  \begin{description}
    \item [Text] Never before had ski racing, a sport dominated by
      monosyllabic mountain men, seen the likes of Alberto Tomba, the
      flamboyant Bolognese flatlander who at 21 captured two gold
      medals at the Calgary Olympics.
    \item [Hypothesis] Alberto Tomba won a race.
  \end{description}
\end{exampleblock}

\end{frame}

\begin{frame}
  \frametitle{Textual Inference}

\begin{exampleblock}{Example 2}
  \begin{description}
    \item [Text] Researchers at the Harvard School of Public Health
      say that people who drink coffee may be doing a lot more than
      keeping themselves awake---this kind of consumption apparently
      also can help reduce the risk of diseases.
    \item [Hypothesis] Coffee drinking has health benefits.
  \end{description}
\end{exampleblock}

\end{frame}

\begin{frame}
  \frametitle{Why is it useful?}

  \begin{description}
    \item [Information Extraction] blah
    \item [Spoken Dialogue] blah
    \item [Question Answering] Who won races at the Calgary Olympics?
    \item [Text Summarization] bkah
  \end{description}

\end{frame}

\subsection{Notes}

\begin{frame}
  \frametitle{Notes}
  \begin{itemize}
    \item QA, Summarization, Dialog (NL understanding; holy grail of NLP)
    \item RTE
    \item Low-level components
    \item ambiguity, why NLP is hard, PP attachment, combinatorial explosion
    \item limitations of NLP, avoiding AI-completeness
    \item statistical vs rule-based
    \item limitations of just throwing more data at a problem
  \end{itemize}
\end{frame}

\pgfdeclareimage[width=8cm]{tag-context}{../images/tag-context}
\begin{frame}
  \frametitle{Example of an Image}
  \pgfuseimage{tag-context}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Case Studies}

\subsection{??}

\begin{frame}
  \frametitle{Notes}
  \begin{itemize}
    \item Text classification
    \item Distributional similarity
    \item Parsing
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Getting Involved}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}