### 2007-09-04 slides/preface.tex
/doc/slides/preface.tex * new introductory slides * moved material from introduction.tex
/nltk/test.portuguese2.doctest
* Portuguese translation by Tiago Tresoldi
  (unfortunately rst.py breaks on non-ascii characters)

/data/corpora/mac_morpho/*
* POS-tagged corpus for Brazilian Portuguese
  contributed by the NILC group from the
  University of Sao Paulo in Sao Carlos

svn/trunk@5335

\documentclass{beamer}             % for slides
% \documentclass[handout]{beamer}    % for handout
\input{beamer}

\title{Introduction to Natural Language Processing}

\author{Steven Bird \and Ewan Klein \and Edward Loper}
\institute{
  University of Melbourne, AUSTRALIA
  \and
  University of Edinburgh, UK
  \and
   University of Pennsylvania, USA
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\section{Preface}

\begin{frame}
  \titlepage
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Overview}

\begin{frame}
  \frametitle{Knowledge and Communication in Language}
  \begin{itemize}
    \item human knowledge, human communication, expressed in language
    \item language technologies: process human language automatically
    \item handheld devices: predictive text, handwriting recognition
    \item web search engines: access to information locked up in text
    \item two facets of the multilingual information society:
    \begin{itemize}
      \item natural human-machine interfaces
      \item access to stored information
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Problem}
  \begin{itemize}
    \item awash with language data
    \item inadequate tools (will this ever change?)
    \item overheads: Perl, Prolog, Java
    \item Natural Language Toolkit (NLTK) as a solution
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{NLTK: What you get...}
  \begin{itemize}
    \item Book
    \item Documentation
    \item FAQ
    \item Installation instructions for Python, NLTK, data
    \item Distributions:
      \textit{Windows, Mac OSX, Unix, data, documentation}
    \item CD-ROM:
      \textit{Python, NLTK, documentation, third-party libraries for numerical
        processing and visualization, instructions}
    \item Mailing lists:\\
      \texttt{nltk-announce}, 
      \texttt{nltk-devel}, 
      \texttt{nltk-users}, 
      \texttt{nltk-portuguese}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{NLTK: Who it is for...}
  \begin{itemize}
    \item people who want to learn how to:
    \begin{itemize}
      \item write programs
      \item to analyze written language
    \end{itemize}
    \item does not presume programming abilities:
    \begin{itemize}
      \item working examples
      \item graded exercises
    \end{itemize}
    \item experienced programmers:
    \begin{itemize}
      \item quickly learn Python (if necessary)
      \item Python features for NLP
      \item NLP algorithms and data structures
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{NLTK: What you will learn...}
  \begin{enumerate}
    \item how to analyze language data
    \item key concepts from linguistic description and analysis
    \item how linguistic knowledge is used in NLP components
    \item data structures and algorithms used in NLP and linguistic data management
    \item standard corpora and their use in formal evaluation
    \item organization of the field of NLP
    \item skills in Python programming for NLP
  \end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{NLTK: Your likely goals...}
\small
\begin{tabular}{p{0.18\textwidth}|p{0.38\textwidth}|p{0.38\textwidth}}
\textbf{Goals} & \multicolumn{2}{c}{\textbf{Background}} \\
& \textit{Arts and Humanities} & \textit{Science and Engineering} \\ \hline
\textbf{Language Analysis}
& Programming to manage language data, explore linguistic models, and test empirical claims
& Language as a source of interesting problems in data modeling, data mining, and knowledge discovery\\ \hline
\textbf{Language Technology}
& Learning to program, with applications to familiar problems, to work in language technology or other technical field
& Knowledge of linguistic algorithms and data structures for high quality, maintainable language processing software\\
\end{tabular}

\end{frame}

\section{NLTK Book}

\begin{frame}
  \frametitle{Philosophy}
  \begin{itemize}
    \item practical
    \item programming
    \item principled
    \item pragmatic
    \item pleasurable
    \item portal    
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Structure}
  \begin{itemize}
    \item Three parts:
    \begin{enumerate}
      \item \textbf{Basics:} text processing, tokenization, tagging,
        lexicons, simple engineering
      \item \textbf{Parsing:} phrase structure, trees, grammars, parsing
      \item \textbf{Advanced Topics:} selected topics in greater depth:
        feature-based grammar, semantics, linguistic data management
    \end{enumerate}
    \item each part: chapter on programming; three chapters on NLP
    \item each chapter: motivation, sections, graded exercises, summary, further reading
  \end{itemize}
\end{frame}

\section{The Python Programming Language}

\subsection{Key Features}

\begin{frame}
  \frametitle{Python: Key Features}
  \begin{itemize}
    \item simple yet powerful, shallow learning curve
    \item object-oriented: encapsulation, re-use
    \item scripting language, facilitates interactive exploration
    \item excellent functionality for processing linguistic data
    \item extensive standard library, incl graphics, web, numerical processing
    \item downloaded for free from \url{http://www.python.org/}
  \end{itemize}
\end{frame}

\subsection{Python Example}

\begin{frame}[fragile]
  \frametitle{Python Example}

\begin{verbatim}
 import sys
 for line in sys.stdin.readlines():
     for word in line.split():
         if word.endswith('ing'):
             print word
\end{verbatim}

\begin{enumerate}
\item whitespace: nesting lines of code; scope
\item object-oriented: attributes, methods (e.g. \texttt{line})
\item readable
\end{enumerate}
\end{frame}

\subsection{Comparison with Perl}
\begin{frame}[fragile]
\frametitle{Comparison with Perl}

\begin{verbatim}
  while (<>) {
      foreach my $word (split) {
          if ($word =~ /ing$/) {
              print "$word\n";
          }
      }
  }
\end{verbatim}

\begin{enumerate}
\item syntax is obscure: \textit{what are:} \verb|<> $ my split| ?
\item ``it is quite easy in Perl to write programs that simply
  look like raving gibberish, even to experienced Perl programmers''
  (Hammond \textit{Perl Programming for Linguists} 2003:47)
\item large programs difficult to maintain, reuse
\end{enumerate}
\end{frame}

\subsection{What NLTK adds to Python}
\begin{frame}
\frametitle{What NLTK adds to Python}

NLTK defines a basic infrastructure that can be used to build NLP
programs in Python.  It provides:

\begin{itemize}
\item Basic classes for representing data relevant to natural language
  processing
\item Standard interfaces for performing tasks, such
  as tokenization, tagging, and parsing
\item Standard implementations for each task, which
  can be combined to solve complex problems
\item Extensive documentation, including tutorials
  and reference documentation
\end{itemize}
\end{frame}

\begin{frame}
  \frametitle{NLTK Design: Requirements}
  \begin{enumerate}
    \item \textbf{simplicity:} intuitive framework with substantial building blocks
    \item \textbf{consistency:} uniform data structures, interfaces --- predictability
    \item \textbf{extensibility:} accommodates new components
      (replicate vs extend exiting functionality)
    \item \textbf{modularity:} interaction between components
    \item \textbf{well-documented:} substantial documentation
  \end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{NLTK Design: Non-requirements}
  \begin{enumerate}
    \item \textbf{encyclopedic:} has many gaps; opportunity for students to extend it
    \item \textbf{efficiency:} not highly optimised for runtime performance
    \item \textbf{programming tricks:} avoid in preference for clear implementations
      (replicate vs extend exiting functionality)
    \item \textbf{modularity:} interaction between components
    \item \textbf{well-documented:} substantial documentation
  \end{enumerate}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Corpora Distributed with NLTK}
  \tiny
  \begin{itemize}
    \item Australian ABC News, 2 genres, 660k words, sentence-segmented
    \item Brown Corpus, 15 genres, 1.15M words, tagged
    \item CMU Pronouncing Dictionary, 127k entries
    \item CoNLL 2000 Chunking Data, 270k words, tagged and chunked
    \item CoNLL 2002 Named Entity, 700k words, pos- and named-entity-tagged (Dutch, Spanish)
    \item Floresta Treebank, 9k sentences (Portuguese)
    \item Genesis Corpus, 6 texts, 200k words, 6 languages
    \item Gutenberg (sel), 14 texts, 1.7M words
    \item Indian POS-Tagged Corpus, 60k words pos-tagged (Bangla, Hindi, Marathi, Telugu)
    \item NIST 1999 Info Extr (sel), 63k words, newswire and named-entity SGML markup
    \item Names Corpus, 8k male and female names
    \item PP Attachment Corpus, 28k prepositional phrases, tagged as noun or verb modifiers
    \item Presidential Addresses, 485k words, formatted text
    \item Roget's Thesaurus, 200k words, formatted text
    \item SEMCOR, 880k words, part-of-speech and sense tagged
    \item SENSEVAL 2, 600k words, part-of-speech and sense tagged
    \item Shakespeare XML Corpus (sel), 8 books
    \item Stopwords Corpus, 2,400 stopwords for 11 languages
    \item Switchboard Corpus (sel), 36 phonecalls, transcribed, parsed
    \item Univ Decl Human Rights, 480k words, 300+ languages
    \item US Pres Addr Corpus, 480k words
    \item Penn Treebank (sel), 40k words, tagged and parsed
    \item TIMIT Corpus (sel), audio files and transcripts for 16 speakers
    \item Wordlist Corpus, 960k words and 20k affixes for 8 languages
    \item WordNet, 145k synonym sets
  \end{itemize}
\end{frame}

\end{document}